To be FAIR: Theory Specification Needs an Update

### 2025-06-19 action_letter.Rmd
update 2025-06-19 22:16:44.792218

<!-- Make comparison to preregistration; that was first proposed and only later evaluated. -->
<!-- Mention how few of preregistered hypotheses are true, and yet how theory is rarely updated? Thus, falsificationism is not working in practice. -->

<!-- <!-- AB: --> -->
<!-- 2.	I think the policies argument is interesting. I agree that currently we simply claim that establish FAIR theories will profit the advancement of science by an argument of analogy (to open data and open code) but ultimately we do not know. For one, I think it's fair to admit this. Second, I would argue that FAIR theories is not a very strict policy (in contrast to other stances in science, which require very specific ways of working, e.g. commiting to falsificationism, adhering to ethical principles, etc.). Really, FAIR theories are no restriction in that sense but more of a protocol for communication, which ultimately does not change your workflow or your methods. -->

@@

We agree with the Editor that there is no guarantee that *"these things are crucial to scientific advance in psychology"*. We have, however, provided reasoned arguments to support the "affordances" of FAIR theory: it *makes possible* particular scientific advancements, even if it does not guarantee them.

Compare FAIR theory to preregistration: it, too, was introduced first as a good idea.
Only now - a decade later - are empirical studies are being conducted to investigate its effects.on scientific practices.

To address this comment, <font color="red">in the Revision:

* We explicitly state that we describe the affordances of FAIR theory and argue for its potential impact by comparing it to other open science methods
* We acknowledge in the Limitations section that empirical evidence for these potential benefits remains to be evaluated.

</font>

<!-- However, our paper is a different kind of paper than an empirical study: we provide a reasoned argument to justify FAIR theory, based on philosophy of science   -->

<!--  Unlike more prescriptive norms—such as commitments to falsificationism or methodological registration—FAIR theory does not constrain a researcher’s workflow or analytical choices. It simply makes theoretical assumptions transparent, findable, and versioned. This distinction is important: the goal is not to impose rigid procedures, but to enable a more cumulative and accountable theoretical discourse. Still, we agree with the reviewer that policies and infrastructure alone are insufficient to drive meaningful change. Culture, norms, and the nature of the subject matter all shape how—and whether—reforms take root. In that spirit, we see FAIR theory not as a top-down imposition, but as a bottom-up opportunity: it offers immediate utility to researchers who want their theories to be discoverable, citable, and improvable. Over time, its adoption may well depend not on mandates, but on whether it proves valuable enough to become a “no-brainer” for the psychological sciences as it has in fields like genetics. -->

>>>>

@@

We agree with the Editor that there is no guarantee that *"these things are crucial to scientific advance in psychology"*. We have, however, provided reasoned arguments to support the *potential impact* of FAIR theory: it makes particular scientific advancements possible. It is up to the scholarly community whether this potential impact will be realized.

To address this comment, in the Revision, we repeatedly refer to the *potential* impact of FAIR theory, thus making it clear that we do not assume that this impact will be automatic nor guaranteed. Furthermore, we acknowledge in the Limitations section that empirical evidence for these potential benefits remains to be evaluated:

`r get_revision("evaluateimpact")`
